---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(modtrees)
```

As a Shiny app grows, the module hierarchy gets harder to hold in your head. You end up tracing through multiple files to answer a simple question: which module calls which? `modtrees` addresses this by statically parsing the R source files in a directory and printing the call graph as a plain-text tree, without running the app.

This vignette walks through the six example applications bundled with `modtrees`. Each one isolates a specific scenario you're likely to encounter when using `mod_tree()` on a real project.

## The `mod_tree()` function

`mod_tree()` takes four arguments:

- `path`: the directory containing R source files (defaults to `"R"`)
- `app_fun`: the top-level launch function name (defaults to `"launch"`)
- `ui_fun`: the UI function name (defaults to `"app_ui"`)
- `server_fun`: the server function name (defaults to `"app_server"`)

The function parses every `.R` file in `path`, identifies functions that use `NS()` or `moduleServer()` as Shiny modules, builds a call graph, and performs a depth-first traversal starting from `app_fun`. The result is printed to the console.

Use `run_demo()` to launch any of the bundled apps interactively so you can inspect their source before running `mod_tree()` on them.

## Default names

The simplest case: a single module nested under `app_ui` and `app_server`, with all function names matching `mod_tree()`'s defaults. No arguments beyond `path` are needed.

```{r eval=FALSE}
run_demo("01-default-names")
```

The `R/` directory for this app contains `launch.R`, `app_ui.R`, `app_server.R`, and `mod_greet.R`. Point `mod_tree()` at it:

```{r eval=FALSE}
mod_tree("inst/apps/01-default-names/R")
```

```verbatim
█─launch
├─█─app_ui
│ └─█─mod_greet_ui
└─█─app_server
  └─█─mod_greet_server
```

The root is `launch`, its two children are `app_ui` and `app_server`, and each carries a single module leaf.

## Custom names

Real apps rarely use `launch`, `app_ui`, and `app_server` verbatim. When the top-level functions have different names, pass them explicitly.

```{r eval=FALSE}
run_demo("02-custom-names")
```

This app uses `run_app()` as the launcher, `ui()` for the interface, and `server()` for the server logic. Calling `mod_tree()` without the extra arguments would start the traversal from a function named `"launch"` that doesn't exist, producing an empty tree. Supply the correct names instead:

```{r eval=FALSE}
mod_tree(
  "inst/apps/02-custom-names/R",
  app_fun    = "run_app",
  ui_fun     = "ui",
  server_fun = "server"
)
```

```verbatim
█─run_app
├─█─ui
│ └─█─mod_counter_ui
└─█─server
  └─█─mod_counter_server
```

The tree is structurally identical to the previous one; only the function names differ.
