---
title: "Inspecting Shiny module structure with modtrees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inspecting Shiny module structure with modtrees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(modtrees)
```

As a Shiny app grows, the module hierarchy gets harder to hold in your head. You end up tracing through multiple files to answer a simple question: which module calls which? `modtrees` addresses this by statically parsing the R source files in a directory and printing the call graph as a plain-text tree, without running the app.

This vignette walks through the six example applications bundled with `modtrees`. Each one isolates a specific scenario you're likely to encounter when using `mod_tree()` on a real project.

## The `mod_tree()` function

`mod_tree()` takes four arguments:

- `path`: the directory containing R source files (defaults to `"R"`)
- `app_fun`: the top-level launch function name (defaults to `"launch"`)
- `ui_fun`: the UI function name (defaults to `"app_ui"`)
- `server_fun`: the server function name (defaults to `"app_server"`)

The function parses every `.R` file in `path`, identifies functions that use `NS()` or `moduleServer()` as Shiny modules, builds a call graph, and performs a depth-first traversal starting from `app_fun`. The result is printed to the console.

Use `run_demo()` to launch any of the bundled apps interactively so you can inspect their source before running `mod_tree()` on them.

## Default names

The simplest case: a single module nested under `app_ui` and `app_server`, with all function names matching `mod_tree()`'s defaults. No arguments beyond `path` are needed.

```{r eval=FALSE}
run_demo("01-default-names")
```

The `R/` directory for this app contains `launch.R`, `app_ui.R`, `app_server.R`, and `mod_greet.R`. Point `mod_tree()` at it:

```{r eval=FALSE}
mod_tree("inst/apps/01-default-names/R")
```

```verbatim
█─launch
├─█─app_ui
│ └─█─mod_greet_ui
└─█─app_server
  └─█─mod_greet_server
```

The root is `launch`, its two children are `app_ui` and `app_server`, and each carries a single module leaf.

## Custom names

Real apps rarely use `launch`, `app_ui`, and `app_server` verbatim. When the top-level functions have different names, pass them explicitly.

```{r eval=FALSE}
run_demo("02-custom-names")
```

This app uses `run_app()` as the launcher, `ui()` for the interface, and `server()` for the server logic. Calling `mod_tree()` without the extra arguments would start the traversal from a function named `"launch"` that doesn't exist, producing an empty tree. Supply the correct names instead:

```{r eval=FALSE}
mod_tree(
  "inst/apps/02-custom-names/R",
  app_fun    = "run_app",
  ui_fun     = "ui",
  server_fun = "server"
)
```

```verbatim
█─run_app
├─█─ui
│ └─█─mod_counter_ui
└─█─server
  └─█─mod_counter_server
```

The tree is structurally identical to the previous one; only the function names differ.

## Nested modules

When a module calls child modules internally, the tree gains depth. `mod_tree()` follows the call graph recursively, so nested relationships appear at the correct level.

```{r eval=FALSE}
run_demo("03-nested-modules")
```

Here `mod_dashboard` is the parent module. Its UI and server each call `mod_chart`, making `mod_chart` a grandchild of the app root:

```{r eval=FALSE}
mod_tree("inst/apps/03-nested-modules/R")
```

```verbatim
█─launch
├─█─app_ui
│ └─█─mod_dashboard_ui
│   └─█─mod_chart_ui
└─█─app_server
  └─█─mod_dashboard_server
    └─█─mod_chart_server
```

The indentation reflects the true call depth. If `mod_chart` itself called further modules, they would appear one level deeper.

## No modules

Not every Shiny app uses modules. When `mod_tree()` finds no functions that call `NS()` or `moduleServer()`, the tree collapses to just the three top-level functions.

```{r eval=FALSE}
run_demo("04-no-modules")
```

```{r eval=FALSE}
mod_tree("inst/apps/04-no-modules/R")
```

```verbatim
█─launch
├─█─app_ui
└─█─app_server
```

A flat tree like this is a useful signal: either the app genuinely has no modules, or the module-detection pass missed something. In the latter case, check that the module functions actually contain `NS()` or `moduleServer()` calls at the top level of their bodies.

## Many modules

As the number of sibling modules grows, the tree expands horizontally. This app wires five modules directly into `app_ui` and `app_server`.

```{r eval=FALSE}
run_demo("05-many-modules")
```

```{r eval=FALSE}
mod_tree("inst/apps/05-many-modules/R")
```

```verbatim
█─launch
├─█─app_ui
│ ├─█─mod_header_ui
│ ├─█─mod_filter_ui
│ ├─█─mod_table_ui
│ ├─█─mod_chart_ui
│ └─█─mod_footer_ui
└─█─app_server
  ├─█─mod_header_server
  ├─█─mod_filter_server
  ├─█─mod_table_server
  ├─█─mod_chart_server
  └─█─mod_footer_server
```

The order of children in each branch follows the order in which the functions are called in the source. `mod_filter` appears second because it is the second module called in `app_ui()`.

## Realistic app with helper functions

The previous examples are deliberately minimal. Production apps include utilities alongside modules: logging helpers, data loaders, plot functions, and other non-module code. This app has all of that.

```{r eval=FALSE}
run_demo("06-realistic-with-helpers")
```

The `R/` directory contains 14 files, including `logr_msg.R`, `scatter_plot.R`, `display_type.R`, and others that define regular functions. These are called from within modules but they are not modules themselves. `mod_tree()` correctly excludes them from the tree. Provide the non-default function names for this app:

```{r eval=FALSE}
mod_tree(
  "inst/apps/06-realistic-with-helpers/R",
  app_fun    = "launch_app",
  ui_fun     = "movies_ui",
  server_fun = "movies_server"
)
```

```verbatim
█─launch_app
├─█─movies_ui
│ ├─█─mod_var_input_ui
│ ├─█─mod_aes_input_ui
│ └─█─mod_scatter_display_ui
└─█─movies_server
  ├─█─mod_var_input_server
  ├─█─mod_aes_input_server
  └─█─mod_scatter_display_server
```

Even though `mod_scatter_display_server` calls `scatter_plot()`, that function does not appear in the tree. `mod_tree()` includes a function in the output only if it is a detected module or one of the three named top-level functions.

## Recap

`mod_tree()` statically parses R source files to build and print a module call graph without running the app. The six scenarios covered here represent the range of structures you are likely to encounter:

1. **Default names** — no arguments beyond `path` are needed when the app follows `launch`/`app_ui`/`app_server` conventions.
2. **Custom names** — supply `app_fun`, `ui_fun`, and `server_fun` when the top-level functions use different names.
3. **Nested modules** — `mod_tree()` follows the call graph recursively, so child modules appear at the correct depth.
4. **No modules** — a flat tree is valid output; it can also indicate a detection miss worth investigating.
5. **Many modules** — sibling modules fan out horizontally, ordered by call sequence in the source.
6. **Helper functions** — regular functions called by modules are excluded from the tree; only `NS()`- or `moduleServer()`-bearing functions are treated as modules.
