---
title: "Modules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(modtrees)
```

This vignette walks through the example Shiny apps using modules.

## Nested modules

When a module calls child modules internally, the tree gains depth. `mod_tree()` follows the call graph recursively, so nested relationships appear at the correct level.

```{r eval=FALSE}
run_demo("03-nested-modules")
```

Here `mod_dashboard` is the parent module. Its UI and server each call `mod_chart`, making `mod_chart` a grandchild of the app root:

```{r eval=FALSE}
mod_tree("inst/apps/03-nested-modules/R")
```

```verbatim
в–Ҳв”Җlaunch
в”ңв”Җв–Ҳв”Җapp_ui
в”Ӯ в””в”Җв–Ҳв”Җmod_dashboard_ui
в”Ӯ   в””в”Җв–Ҳв”Җmod_chart_ui
в””в”Җв–Ҳв”Җapp_server
  в””в”Җв–Ҳв”Җmod_dashboard_server
    в””в”Җв–Ҳв”Җmod_chart_server
```

The indentation reflects the true call depth. If `mod_chart` itself called further modules, they would appear one level deeper.

## No modules

Not every Shiny app uses modules. When `mod_tree()` finds no functions that call `NS()` or `moduleServer()`, the tree collapses to just the three top-level functions.

```{r eval=FALSE}
run_demo("04-no-modules")
```

```{r eval=FALSE}
mod_tree("inst/apps/04-no-modules/R")
```

```verbatim
в–Ҳв”Җlaunch
в”ңв”Җв–Ҳв”Җapp_ui
в””в”Җв–Ҳв”Җapp_server
```

A flat tree like this is a useful signal: either the app genuinely has no modules, or the module-detection pass missed something. In the latter case, check that the module functions actually contain `NS()` or `moduleServer()` calls at the top level of their bodies.

## Many modules

As the number of sibling modules grows, the tree expands horizontally. This app wires five modules directly into `app_ui` and `app_server`.

```{r eval=FALSE}
run_demo("05-many-modules")
```

```{r eval=FALSE}
mod_tree("inst/apps/05-many-modules/R")
```

```verbatim
в–Ҳв”Җlaunch
в”ңв”Җв–Ҳв”Җapp_ui
в”Ӯ в”ңв”Җв–Ҳв”Җmod_header_ui
в”Ӯ в”ңв”Җв–Ҳв”Җmod_filter_ui
в”Ӯ в”ңв”Җв–Ҳв”Җmod_table_ui
в”Ӯ в”ңв”Җв–Ҳв”Җmod_chart_ui
в”Ӯ в””в”Җв–Ҳв”Җmod_footer_ui
в””в”Җв–Ҳв”Җapp_server
  в”ңв”Җв–Ҳв”Җmod_header_server
  в”ңв”Җв–Ҳв”Җmod_filter_server
  в”ңв”Җв–Ҳв”Җmod_table_server
  в”ңв”Җв–Ҳв”Җmod_chart_server
  в””в”Җв–Ҳв”Җmod_footer_server
```

The order of children in each branch follows the order in which the functions are called in the source. `mod_filter` appears second because it is the second module called in `app_ui()`.

## Using `mod_tree_app()` on a live object

`mod_tree()` reads files from disk. `mod_tree_app()` inspects a `shiny.appobj` in memory instead; it recovers the server function directly from the app object and identifies modules by walking the same environment. The result is the same tree.

Using the five-module app as an example, build the app object by passing function references (not evaluated calls):

```{r eval=FALSE}
app <- shiny::shinyApp(ui = app_ui, server = app_server)
```

`mod_tree_app()` requires no `path` argument. The module structure comes from the server's enclosing environment, so the call is shorter when function names match the defaults:

```{r eval=FALSE}
mod_tree_app(app)
```

```verbatim
в–Ҳв”Җapp
в”ңв”Җв–Ҳв”Җapp_ui
в”Ӯ в”ңв”Җв–Ҳв”Җmod_header_ui
в”Ӯ в”ңв”Җв–Ҳв”Җmod_filter_ui
в”Ӯ в”ңв”Җв–Ҳв”Җmod_table_ui
в”Ӯ в”ңв”Җв–Ҳв”Җmod_chart_ui
в”Ӯ в””в”Җв–Ҳв”Җmod_footer_ui
в””в”Җв–Ҳв”Җapp_server
  в”ңв”Җв–Ҳв”Җmod_header_server
  в”ңв”Җв–Ҳв”Җmod_filter_server
  в”ңв”Җв–Ҳв”Җmod_table_server
  в”ңв”Җв–Ҳв”Җmod_chart_server
  в””в”Җв–Ҳв”Җmod_footer_server
```

The root label defaults to `"app"` because the `shiny.appobj` does not record the name of the launch function. Pass `app_fun` to override it:

```{r eval=FALSE}
mod_tree_app(app, app_fun = "launch")
```

The tree structure is unchanged; only the root label differs.
