---
title: "Test coverage and traceability"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Test coverage and traceability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

```{r setup}
library(modtrees)
```

This vignette records which tests cover which functions, summarizes the current state of the test suite, and documents known issues. It is intended as a living document: update the matrix whenever tests are added or functions change.

## Functions

`modtrees` exposes two exported functions and five internal helpers.

| Function | Type | Test file |
|---|---|---|
| `mod_tree()` | exported | `test-mod_tree.R` |
| `mod_tree_app()` | exported | `test-mod_tree_app.R` |
| `run_demo()` | exported | — |
| `build_tree()` | internal | `test-build_tree.R` |
| `extract_func_def()` | internal | `test-extract_func_def.R` |
| `find_calls()` | internal | `test-find_calls.R` |
| `render_tree()` | internal | `test-render_tree.R` |
| `uses_ns()` | internal | `test-uses_ns.R` |

`run_demo()` is an interactive launcher with no automated tests. Its correctness is verified manually by running each bundled app.

## Traceability matrix

Columns mark every function exercised by each test. Integration-level tests (those in `test-mod_tree.R` and `test-mod_tree_app.R`) exercise multiple internal helpers in a single call; those cells are marked accordingly.

```{r matrix, echo=FALSE}
tests <- data.frame(
  Description = c(
    # build_tree
    "builds simple tree",
    "handles cycles",
    "handles nested modules",
    "unknown root produces a root node with no children",
    # extract_func_def
    "extracts <- function assignment",
    "extracts = function assignment",
    "returns NULL for non-function assignment",
    "returns NULL for bare expressions",
    "extracts assign() style",
    # find_calls
    "finds direct function calls",
    "finds bare name references",
    "ignores unknown names",
    "returns unique results",
    "finds calls nested inside control flow",
    # mod_tree_app
    "errors on non-shiny.appobj input",
    "detects one module when UI is an unevaluated function",
    "app_fun sets the root label",
    "auto-detects UI function name when passed unevaluated",
    "ui_fun fallback works when UI is pre-evaluated",
    "flat tree when app has no modules",
    "nested modules appear at the correct depth",
    "multiple sibling modules all appear under server",
    "non-module helpers called by modules do not appear in tree",
    "invisibly returns the tree list",
    # mod_tree
    "mod_tree works on a minimal app",
    "mod_tree errors gracefully on empty directory",
    "custom app_fun, ui_fun, server_fun are honoured",
    "nested modules appear at the correct depth (file-based)",
    "no modules produces a flat tree with only top-level functions",
    # render_tree
    "renders root node",
    "renders single child",
    "renders multiple children with correct connectors",
    "grandchild indented correctly with continuation line",
    # uses_ns
    "detects direct NS() call",
    "detects shiny::NS() call",
    "detects moduleServer() call",
    "detects shiny::moduleServer() call",
    "returns FALSE for non-module function"
  ),
  File = c(
    rep("test-build_tree", 4),
    rep("test-extract_func_def", 5),
    rep("test-find_calls", 5),
    rep("test-mod_tree_app", 10),
    rep("test-mod_tree", 5),
    rep("test-render_tree", 4),
    rep("test-uses_ns", 5)
  ),
  mod_tree      = c(rep("",24), rep("\u2713",5), rep("",9)),
  mod_tree_app  = c(rep("",14), rep("\u2713",10), rep("",14)),
  build_tree    = c(rep("\u2713",4), rep("",10),
                    "", "\u2713","\u2713","","\u2713","\u2713","\u2713","\u2713","","",
                    rep("\u2713",5), rep("",9)),
  extract_func_def = c(rep("",4), rep("\u2713",5), rep("",10),
                       rep("\u2713",5), rep("",14)),
  find_calls    = c(rep("",9), rep("\u2713",5),
                    "", "\u2713","","","\u2713","","","","\u2713","",
                    rep("\u2713",5), rep("",9)),
  render_tree   = c(rep("",14),
                    "", "\u2713","\u2713","","","","","","","",
                    rep("\u2713",5), rep("\u2713",4), rep("",5)),
  uses_ns       = c(rep("",14),
                    "", "\u2713","","\u2713","\u2713","\u2713","\u2713","\u2713","\u2713","",
                    rep("\u2713",5), rep("",9)),
  stringsAsFactors = FALSE
)

knitr::kable(
  tests,
  col.names = c("Test description", "File",
                "mod_tree", "mod_tree_app",
                "build_tree", "extract_func_def",
                "find_calls", "render_tree", "uses_ns"),
  align = c("l", "l", rep("c", 7))
)
```

## Coverage notes

`run_demo()` has no automated tests. It wraps `shiny::shinyAppDir()` with path-resolution logic; its correctness depends on the bundled apps in `inst/apps/` being valid, which is implicitly verified whenever those apps are rendered for manual inspection.

The `mod_tree()` integration tests in `test-mod_tree.R` exercise all five internal helpers in combination. A failure there that does not also fail the corresponding unit test points to an interaction bug rather than a bug in the helper itself.

## Known issues

### `extract_func_def()`: vector coercion with complex expressions

Found during wild testing against the `pffoundations` package (July 2026, 35 R source files, ~20 modules).

**Symptom:** `mod_tree()` crashes with:

```
'length = 3' in coercion to 'logical(1)'
Calls: mod_tree -> extract_func_def
```

**Root cause.** Line 17 of `R/extract_func_def.R`:

```r
op <- as.character(expr[[1]])
```

When `expr[[1]]` is a complex call rather than a name, `as.character()` returns a character *vector*, not a scalar. The `%in%` check on line 19 then returns a logical vector, and `&&` cannot coerce it to a scalar.

**Fix.** Take only the first element before the comparison:

```r
op <- as.character(expr[[1]])[1]
if (op %in% c("<-", "=", "assign") && length(expr) == 3) { ...
```

Apply the same `[1]` guard to line 28:

```r
if (is.call(val) && as.character(val[[1]])[1] == "function") {
```

**Tests to add.** A test covering a complex left-hand-side expression that currently causes the crash:

```r
test_that("returns NULL safely when expr[[1]] is a complex call", {
  # e.g. a namespaced assignment: pkg::name <- function() {}
  expr <- parse(text = "a::b <- function() {}")[[1]]
  expect_null(extract_func_def(expr))
})
```

**Status:** Documented; fix and regression test pending.

## Recap

- 38 tests span 7 functions across 6 test files.
- `mod_tree()` and `mod_tree_app()` each have dedicated unit tests plus integration coverage through the internal helpers they call.
- `run_demo()` is untested automatically; manual verification through `inst/apps/` demos covers it.
- One known bug in `extract_func_def()` causes a crash on complex left-hand-side expressions; the fix is a one-character change on two lines.
